MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-24 chapters/MetaTracts.tex
corrections intro + section 2

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The material of choice of a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing  parts made of conventional materials such as aluminum or steel. CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.

In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength, high corrosion resistance. CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
At the same time highly complex and integrated components which were previously  impossible to manufacture may be produced from CFRP's.
For example even primary structures and highly loaded components in aeronautics are being increasingly manufactured from CFRP.

Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components.
A matrix on the one side, which acts as a bonding component and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. 
The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On the one hand ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component.

Industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly being applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. In industrial XCT, the typical scanning geometry places the specimen on a rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays get attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below $500 nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work, we focus on datasets with larger viewports and lower resolutions where the single carbon fibers (filaments) are not or are  barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of  individual fibers.
%of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}. On the other hand, recent developments of industrial 3D X-Ray computed tomography (XCT) with regard to larger detectors, larger field of views and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.

@@

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts, ``geometric structure" and ``spatial context". Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.

The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data. Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provides details of our approach. Section \ref{sec:results} details experiments, section~\ref{sec:user_eval} summarizes our user evaluation results. Section~\ref{sec:param_choices} explains our  parameter choices. We end with conclusions in section~\ref{sec:conclusions}.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\caption{Data Characteristics. (A) shows a rendering of the data. (B) shows 2D slices along the X and (C) along the Z axis. (D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the data type is unsigned 8 bit integer. Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis. Figure~\ref{fig:data-char}(D) shows the magnified version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

>>>>

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The materials of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing  parts made of conventional materials such as aluminum or steel. CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.

In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength, high corrosion resistance. CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
At the same time highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates, consist of two main components:
A matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles. 
The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component.

Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. Typical scanning geometry in industrial XCT is the specimen placed on a rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays get attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or are barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of individual fibers.
%of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}. On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.

@@

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.

The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  section \ref{sec:char_data} describes the   characteristics of our data. Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provide details of our approach. Section \ref{sec:results} details experiments and section~\ref{sec:user_eval} summarizes our user evaluation results. Section~\ref{sec:param_choices} explains our  parameter choices. We end with conclusions in section~\ref{sec:conclusions}.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\caption{Data characteristics: (A) shows a rendering of the data. (B) shows 2D slices along the X- and (C) along the Z-axis. (D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}

Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$, the data type is unsigned 8 bit integer. Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset. Figure~\ref{fig:data-char}(B) shows 2D slices along the $X$- and Figure~\ref{fig:data-char}(C) in $Z$-axis. Figure~\ref{fig:data-char}(D) shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}
